---
title: "Hypothesis Testing 102"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, inclue=FALSE, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, cache = TRUE)
# options(digits = 4)
library(dplyr)
```

# Setup

```{r}
x_bar <- 200
hypo <- 195
s <- 6
n <- 10

```

# Hypothesis testing using test statistic

What we called "Method #3" in class and what we usually think about when we talk about hypothesis testing.

```{r}
signal <- x_bar - hypo
noise <- s/sqrt(n) # SEM

test_statistic <- signal/noise
test_statistic
```

What's the critical value (aka the cutoff) for

According to the Standard Normal Distribution:

```{r}
qnorm(0.975, mean = 0, sd = 1) # 1.96
# qnorm(0.5, mean = 0, sd = 1) # 0
# pnorm(1.96, mean = 0, sd = 1)
```

According to the t Distribution:

```{r}
# The t-dist:
qt(0.975, df = n-1) # 2.26
# qt(0.975, df = 74) # 1.99
```

The p-value:

```{r}
# The 2-sided p-value
(1 - pt(test_statistic, df = n - 1)) * 2
```

# Hypothesis testing using 95% CI

What we called "Method #1" in class and a perfectly acceptable form of hypothesis testing.

```{r}
err_N <- qnorm(0.975, mean = 0, sd = 1) # 1.96
lower_N <- x_bar - (err_N * noise)
upper_N <- x_bar + (err_N * noise)
```

Using the Normal distribution: `r c(lower_N, upper_N)`.

```{r}
err_t <- qt(0.975, df = n - 1) # 2.26
lower_t <- x_bar - (err_t * noise)
upper_t <- x_bar + (err_t * noise)
```

Using the t-distribution: `r c(lower_t, upper_t)`.

# Built in

```{r}

# getwd()
read.delim("~/Documents/Misk_DSI_data/data/martians.txt") %>%
  filter(Site == "Site I") %>%
  select(Height) %>%
  pull(Height) -> site_I_height

t.test(site_I_height, mu = hypo)

```