---
title: "LM vs t-test vs ANOVA"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, inclue=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
options(digits = 2)
library(car)
library(ggplot2)
library(dplyr)
library(tidyr)
```

## 2-sample t-test

The data

```{r}
my_sleep <- sleep[-3]
glimpse(my_sleep)
```

```{r}
summary(my_sleep)
```

```{r}
ggplot(my_sleep, aes(x = group, y = extra)) +
  geom_jitter(width = 0.1) +
  stat_summary(fun.data = mean_sdl, fun.args = list(mult = 1), color = "#e41a1c")
```


1. Do a classic 2-sample t-test

```{r}
# classic 2- sample t-test
t.test(extra ~ group, data = my_sleep, var.equal = TRUE)

```

2. A Welch's version of the 2-sample t-test (don't assume variances are equal)

```{r}
# Welch's 2-sample t-test
t.test(extra ~ group, data = my_sleep)
```

3. As a linear model, using `lm()`. Interpret the results.

```{r}
# As a linear model
my_sleep_lm <- lm(extra ~ group, data = my_sleep)
my_sleep_lm
```

```{r}
summary(my_sleep_lm)
```

4. As an ANOVA, using `anova()`. Interpret the results.

```{r}
# As an ANOVA
anova(my_sleep_lm)

```

## Linear Model

The data:

```{r}
lm_data_wide <- data.frame(xx = c(2, 3, 5, 6),
                      yy = c(2, 5, 4, 7))
glimpse(lm_data_wide)
```

```{r}
ggplot(lm_data_wide, aes(x = xx, y = yy)) +
  geom_point()
```

```{r}
# We should actually have "tidy" data
# MEASURE VARIABLES = columns that should become "tidy"
# here, it's all of them.
lm_data_wide %>% 
  pivot_longer(everything(), 
               names_to = "group", 
               values_to = "value") -> lm_data_tidy
glimpse(lm_data_tidy)
```

```{r}
ggplot(lm_data_tidy, aes(x = group, y = value)) +
  geom_point()

```

1. Use the wide format data and do an `lm()`

```{r}
# An OLS LM
my_lm <- lm(value ~ group, data = lm_data_tidy)
summary(my_lm)
```

2. Can you do a t-test on either of these formats?

Using the wide format:

```{r}
# A t-test
t.test(lm_data_wide$xx, lm_data_wide$yy)
```
Using the tidy format:

```{r}
# A t-test
t.test(value ~ group, data = lm_data_tidy)

```


3. Calculate an anova, using `anova()`.

```{r}
# As an ANOVA
anova(my_lm)
```


## Data for 1-way ANOVA

```{r}
data(PlantGrowth)
glimpse(PlantGrowth)
```

```{r}
summary(PlantGrowth)
```
```{r}
ggplot(PlantGrowth, aes(x = group, y = weight)) +
  geom_jitter(width = 0.1)
```

1. Do an `lm()` on this data.

```{r}
# lm
plant_lm <- lm(weight ~ group, data = PlantGrowth)
summary(plant_lm)
```

2. Calcualte an anova table with `anova()`.

```{r}
# ANOVA
anova(plant_lm)
```

